---
title: "R Notebook"
output: html_notebook
---

```{r}
library(dplyr)
```

```{r}
# Load packages used in this guide ----
packages <- c("tidyverse", "knitr", "kableExtra", "leaps",
              "car", "psych", "DescTools", "pROC", "yardstick", "separationplot")

invisible(
  lapply( 
    X = packages,
    FUN = library,
    character.only = TRUE,
    quietly = TRUE
  )
)

# Set Table Option ----
options(knitr.kable.NA = "") 
```

```{r}
#urlfle <- "https://github.com/apurva-sista/final_stat380/blob/main/dataset-of-00s%20(1).csv"
df1 <- read.csv("dataset-of-00s (1).csv")
df2 <- read.csv("dataset-of-10s (1).csv")

```

```{r}

df3 <- rbind(df1,df2)
```

EDA:

```{r}
head(df3)
```

```{r}
str(df3)
```

```{r}
library(dplyr)

df3_dropped <- dplyr::select(df3, -artist, -uri, -track)

# Display the first few rows of the modified dataframe
head(df3_dropped)
```

```{r}

sum(is.na(df3))
```

```{r}
calculate_stats <- function(x) {
  if (is.numeric(x)) {
    c(Count = length(x), Mean = mean(x), Std = sd(x), Max = max(x), Min = min(x))
  } else {
    NULL
  }
}

# Apply the function to each column of the dataset
stats <- sapply(df3_dropped, calculate_stats)

# Display the statistics
stats
```

```{r}
# Load necessary libraries
library(corrplot)

# Assuming df3 is your dataframe
# Select only the columns that are numeric
df3_numeric <- df3[sapply(df3, is.numeric)]

# Compute the correlation matrix for the numeric columns
cor_matrix <- cor(df3_numeric, use = "complete.obs") # 'use' argument handles missing values

# Visualize the correlation matrix using corrplot
corrplot(cor_matrix, method = "circle")


```

```{r}
# Load necessary libraries
library(ggplot2)
library(reshape2) # for melting the correlation matrix into a long format

# Assuming df3 is your data frame after dropping non-numeric columns
# Calculate Spearman correlation matrix
spearman_correlation <- cor(df3_numeric, method = "spearman")

# Melt the correlation matrix into a long format
melted_corr <- melt(spearman_correlation)

# Create the correlation plot with adjustments for readability
correlation_plot <- ggplot(melted_corr, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") + 
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Spearman\nCorrelation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 5),
        axis.text.y = element_text(size = 5),
        axis.title = element_blank(),
        legend.position = "right") +
  coord_fixed()

# Display the plot in a larger window in RStudio
print(correlation_plot)

```

Since the correlation values between any two variables is not greater than 90%, none of the attributes were removed.

```{r}
# Load necessary libraries
library(tidyverse)

# List of features to plot
features <- c('danceability', 'energy', 'loudness', 'speechiness', 'acousticness', 'instrumentalness', 'valence', 'tempo', 'duration_ms')

# Create a function to plot each feature
plot_feature <- function(feature) {
  ggplot(df3, aes_string(x=feature)) +
    geom_histogram(bins=30, fill="blue", color="black", alpha=0.7) +
    geom_density(color="red") +
    labs(title=paste('Distribution of', feature), x=feature, y='Frequency')
}

# Plot each feature in RStudio
for (feature in features) {
  print(plot_feature(feature))
}

```

From the above cells, it can be concluded that hit songs have high values of loudness, tempo, valence, danceability, energy, time_signature, mode and low values of speechiness, liveness, acousticness, instrumentalness, sections, chorus_hit, duration_ms.

```{r}
modelData <- df3 %>%
  drop_na() %>%
  mutate(
    tempID = row_number()
  )

set.seed(380)
trainingData <- modelData %>%
  slice_sample(prop = 0.8)

trainingResults <- trainingData

testingData <- modelData %>%
  filter(!(tempID %in% trainingData$tempID))

```

```{r}
# General Structure of glm call ----
model1 <- glm(
  formula = target ~ energy,
  data = trainingData,
  family = binomial,
  na.action = "na.omit"
)
```

```{r}

# Form Candidate Model 2 ----
## Lower bound
### Intercept only
lower <- glm(
  formula = target ~ 1,
  data = trainingData,
  family = binomial
)
## Upper bound
### All quantities crossed with species plus island
upper <- glm(
  formula = target ~ (acousticness + instrumentalness + duration_ms + chorus_hit + liveness),
  data = trainingData,
  family = binomial
)

## Stepwise search
model2 <- step(
  object = lower,
  scope = list(
    lower = lower,
    upper = upper
  ),
  data = trainingData,
  direction = "both",
  k = 2,
  trace = 0
)
```

```{r}
# Model 1 Coefficient Table ----
as.data.frame(summary(model1)$coefficients) %>%
  rownames_to_column(var = "term") %>%
  rename(coefficient = Estimate) %>% 
  mutate(
    prob_odds = case_when(
      coefficient == "(Intercept)" ~ exp(coefficient)/(1 + exp(coefficient)),
      .default = exp(coefficient)
    ),
    .after = coefficient
  ) %>%
  kable(
    digits = 3,
    booktabs = TRUE,
    align = c("l", rep("c", 5)),
    col.names = c("Term", "Coefficient", "Prob./Odds Ratio",
                  "Std. Err.", "Z", "p-value"),
    table.attr = 'data-quarto-disable-processing="true"'
  ) %>%
  kable_classic(
    position = "center",
    bootstrap_options = "condensed",
    latex_options = c("HOLD_position"),
    full_width = FALSE
  )
```

```{r}
model1CI <- confint(
  object = model1,
  parm = "energy",
  level = 0.9
)

# Stored fitted values for Model 1 ----
trainingResults$model1Pred <- predict(
  object = model1,
  newdata = trainingData,
  type = "response"
)

# Apply naïve rule ----
trainingResults <- trainingResults %>%
  mutate(
    model1Class = case_when(
      model1Pred > 0.5 ~ 1,
      .default = 0
    )
  )

```

```{r}

library(janitor)
library(knitr)
library(kableExtra)


# Build Confusion Matrix for Model 1 ----
trainingResults %>%
  tabyl(var1 = model1Class, var2 = target) %>%
  adorn_title(
    placement = "combined",
    row_name = "Predicted",
    col_name = "Actual"
  ) %>%
  kable(
    booktabs = TRUE,
    align = "c",
    table.attr = 'data-quarto-disable-processing="true"'
  ) %>%
  kable_classic(
    position = "center",
    bootstrap_options = "condensed",
    latex_options = c("HOLD_position"),
    full_width = FALSE
  )

```

\<\<\<\<\<\<\< HEAD

```{r}
as.data.frame(summary(model2)$coefficients) %>%
  rownames_to_column(var = "term") %>%
  rename(coefficient = Estimate) %>% 
  mutate(
    prob_odds = case_when(
      coefficient == "(Intercept)" ~ exp(coefficient)/(1 + exp(coefficient)),
      .default = exp(coefficient)
    ),
    .after = coefficient
  ) %>%
  mutate(
    `Pr(>|z|)` = ifelse(
      test = `Pr(>|z|)` < 0.001,
      yes = paste("< 0.001"),
      no = round(`Pr(>|z|)`, 3)
    ),
    term = case_when(
      term == "(Intercept)" ~ "Intercept",
      grepl(x = term, pattern = "instrumentalness") ~ "instrumentalness",
      grepl(x = term, pattern = "acousticness") ~ "acousticness",
      grepl(x = term, pattern = "duration_ms") ~ "duration_ms",
      .default = term
    )
  ) %>%
  kable(
    digits = 3,
    booktabs = TRUE,
    align = c("l", rep("c", 5)),
    col.names = c("Term", "Coefficient", "Prob./Odds Ratio",
                  "Std. Err.", "Z", "p-value"),
    table.attr = 'data-quarto-disable-processing="true"'
  ) %>%
  kable_classic(
    position = "center",
    bootstrap_options = "condensed",
    latex_options = c("HOLD_position"),
    full_width = FALSE
  )
```

```{r}
ggplot(
  data = data.frame(
    residuals = residuals(model2, type = "pearson"),
    fitted = fitted(model2)
  ),
  mapping = aes(x = fitted, y = residuals)
) +
  geom_point() +
  geom_smooth(
    formula = y ~ x,
    method = stats::loess,
    method.args = list(degree = 1),
    se = FALSE,
    linewidth = 0.5
  ) +
  theme_bw() +
  labs(
    x = "Fitted",
    y = "Pearson Residuals"
  ) 
```

```{r}
summary_df <- as.data.frame(summary(model2)$coefficients) %>%
rownames_to_column(var = "term") %>%
rename(coefficient = Estimate) %>%
mutate(
prob_odds = case_when(
coefficient == "(Intercept)" ~ exp(coefficient)/(1 + exp(coefficient)),
.default = exp(coefficient)
),
.after = coefficient
) %>%
mutate(
`Pr(>|z|)` = ifelse(
test = `Pr(>|z|)` < 0.001,
yes = paste("< 0.001"),
no = round(`Pr(>|z|)`, 3)
),
term = case_when(
term == "(Intercept)" ~ "Intercept",
      grepl(x = term, pattern = "instrumentalness") ~ "instrumentalness",
      grepl(x = term, pattern = "acousticness") ~ "acousticness",
      grepl(x = term, pattern = "duration_ms") ~ "duration_ms",
.default = term
)
)

```

```{r}
trainingResults$model2Pred <- predict(
  object = model2,
  newdata = trainingData,
  type = "response"
)

trainingResults <- trainingResults %>%
  mutate(
    model2Class = case_when(
      model2Pred > 0.5 ~ 1,
      .default = 0
    )
  )
```

```{r}

trainingResults %>%
  tabyl(var1 = model2Class, var2 = target) %>%
  adorn_title(
    placement = "combined",
    row_name = "Predicted",
    col_name = "Actual"
  ) %>%
  kable(
    booktabs = TRUE,
    align = "c",
    table.attr = 'data-quarto-disable-processing="true"'
  ) %>%
  kable_classic(
    position = "center",
    bootstrap_options = "condensed",
    latex_options = c("HOLD_position"),
    full_width = FALSE
  )

```

```{r}
model1ROC <- roc(
  formula = target ~ model1Pred,
  data = trainingResults
)
model1ROC_df <- data.frame(
  threshold = model1ROC$thresholds,
  sensitivity = model1ROC$sensitivities,
  specificity = model1ROC$specificities,
  model = "Model 1"
)

```

```{r}
model2ROC <- roc(
  formula = target ~ model2Pred,
  data = trainingResults
)
model2ROC_df <- data.frame(
  threshold = model2ROC$thresholds,
  sensitivity = model2ROC$sensitivities,
  specificity = model2ROC$specificities,
  model = "Model 2"
)

```

```{r}
rocData <- rbind(model1ROC_df, model2ROC_df)

```

```{r}
aucData <- data.frame(
  model = c("Model 1", "Model 2"),
  auc = c(model1ROC$auc, model2ROC$auc)
)
```

```{r}


ggplot(
  data = rocData,
  mapping = aes(x = 1 - specificity, y = sensitivity, color = model)
) +
  geom_path() +
  geom_abline(
    slope = 1,
    intercept = 0,
    linetype = "dotted"
  ) +
  geom_text(
    inherit.aes = FALSE,
    data = aucData,
    mapping = aes(label = paste(model, "AUC: \n", round(auc, 3))),
    x = c(0.25, 0.25),
    y = c(0.4, 0.9)
  ) +
  theme_bw() +
  coord_fixed()
```

```{r}

trainingResults <- trainingResults %>%
  mutate(
    targetNum = case_when(
      target == 0 ~ 0,
      target == 1 ~ 1
    ),
    .after = target
  ) # separationplot requies actual to be numeric 0/1
par(mfrow = c(1, 2), mar = c(4,0,4,0))

```

```{r}
separationplot(
  pred = trainingResults$model1Pred, 
  actual = trainingResults$targetNum, 
  type = "line", 
  line = TRUE, 
  show.expected = TRUE, 
  heading = "Model 1",
  newplot = FALSE
)

```

```{r}
## Model 2
separationplot(
  pred = trainingResults$model2Pred, 
  actual = trainingResults$targetNum, 
  type = "line", 
  line = TRUE, 
  show.expected = TRUE, 
  heading = "Model 2",
  newplot = FALSE
)
```

```{r}
testingData <- testingData %>%
  mutate(
    targetNum = case_when(
      target == 0 ~ 0,
      target == 1 ~ 1
    ),
    .after = target
  )
testingData$predict <- predict(
  object = model2,
  newdata = testingData,
  type = "response"
)
testingData <- testingData %>%
  mutate(
    model2Class = case_when(
      predict > 0.5 ~ 1,
      .default = 0
    )
  )

```

```{r}
testingData %>%
  tabyl(var1 = model2Class, var2 = target) %>%
  adorn_title(
    placement = "combined",
    row_name = "Predicted",
    col_name = "Actual"
  ) %>%
  kable(
    booktabs = TRUE,
    align = "c",
    table.attr = 'data-quarto-disable-processing="true"'
  ) %>%
  kable_classic(
    position = "center",
    bootstrap_options = "condensed",
    latex_options = c("HOLD_position"),
    full_width = FALSE
  )

```



DECISION TREES FROM HERE

```{r}
packages <- c("tidyverse", "knitr", "kableExtra", "janitor",
              "tree", "rpart", "rpart.plot", "partykit",
              "rattle", "randomForest", "yardstick")
```


```{r}
invisible(
  lapply( 
    X = packages,
    FUN = library,
    character.only = TRUE,
    quietly = TRUE
  )
)
```



```{r}
df3$target <- as.factor(df3$target)

treeModel <- df3 %>%
  dplyr::select(!energy) %>%
  drop_na() %>%
  mutate(
    tempID = row_number(),
    .before = uri
  )
```

```{r}
set.seed(380)
treeTrainingData <- treeModel %>%
  group_by(danceability, key, target) %>%
  slice_sample(prop = 0.8)
```

```{r}
treeTestingData <- treeModel %>%
  filter(!(track %in% treeTrainingData$track))
```


```{r}
# Grow penguin tree via rpart package
# library(rpart)
rpartSong1 <- rpart(
  formula = target ~ loudness + speechiness,
  data = treeTrainingData,
  method = "class",
  parms = list(split = "information")
  # control = rpart.control()
)
```


```{r}
# Display rpart.plot ----
# library(rpart.plot)
rpart.plot(
  x = rpartSong1,
  type = 2,
  extra = 101
)
```



```{r}
# Get table elements ----
invisible(capture.output({cpTable <- printcp(rpartSong1)}))

# Create nice looking table of CP results ----
kable(
  x = cpTable,
  col.names = c("CP", "Num. of splits", "Rel. Error",
                "Mean Error", "Std. Deviation of Error"),
  digits = 3,
  booktabs = TRUE,
  align = "c",
  table.attr = 'data-quarto-disable-processing="true"'
) %>%
  kable_classic(
    full_width = FALSE
  )
```



```{r}
# Plot the CP results from rpart ----
plotcp(
  x = rpartSong1,
  minline = TRUE,
  upper = "size"
)
```


```{r}
# Prune the rpart Penguin Tree ----
rpartSong2 <- prune(
  tree = rpartSong1,
  cp = 0.019
)
```


```{r}
# Display rpart.plot ----
# library(rpart.plot)
rpart.plot(
  x = rpartSong2,
  type = 2,
  extra = 101
)
```



```{r}
# Using the predict function on the testing data ----
## The rpart package
pred_songsRpart2 <- predict(
  object = rpartSong2,
  newdata = treeTestingData,
  type = "prob"
)
```


```{r}
pred_songsRpart2
```

```{r}
# Data Wrangling the predictions ----
songPredictions <- data.frame(
  rpart2_female = pred_songsRpart2[, 1],
  rpart2_male = pred_songsRpart2[, 2]
) %>%
  mutate(
    rpart2_pred = ifelse(
      test = rpart2_female > rpart2_male,
      yes = "female",
      no = "male"
    )
  )

## Set predictions as factors
songPredictions$rpart2_pred <- as.factor(songPredictions$rpart2_pred)

# Merge supervision column into predictions data frame ----
songPredictions <- cbind(
  tempID = treeTestingData$tempID,
  target = treeTestingData$target,
  songPredictions
)
```


```{r}
songPredictions$target <- ifelse(songPredictions$target == 0, "female", "male")
songPredictions$target <- as.factor(songPredictions$target)
```

```{r}
# Build confusion matrix for second tree model ----
# library(yardstick)
conf_mat(
  data = songPredictions,
  truth = target,
  estimate = rpart2_pred
)$table %>%
  kable(
    col.names = c("Prediction/Supervision", "Female", "Male"),
  digits = 3,
  booktabs = TRUE,
  # caption = "Your caption here",
  align = "c",
  table.attr = 'data-quarto-disable-processing="true"'
  ) %>%
  kable_classic(
    full_width = FALSE
  )
```


```{r}
# Build a data frame with model metrics ----
songPreds <- songPredictions %>%
  dplyr::select(tempID, target, contains("_pred")) %>%
  pivot_longer(
    cols = !c(tempID, target),
    names_to = "model",
    values_to = "prediction"
  )

accuracy <- songPreds %>%
  group_by(model) %>%
  accuracy(
    truth = target,
    estimate = prediction
  )

sensitivity <- songPreds %>%
  group_by(model) %>%
  sensitivity(
    truth = target,
    estimate = prediction,
    event_level = "second"
  )

specificity <- songPreds %>%
  group_by(model) %>%
  specificity(
    truth = target,
    estimate = prediction,
    event_level = "second"
  )

modelMetrics <- bind_rows(
  accuracy,
  sensitivity,
  specificity
)
```


```{r}
# Make a nice looking table of model metrics ----
modelMetrics %>%
  dplyr::select(model, .metric, .estimate) %>%
  pivot_wider(
    id_cols = model,
    names_from = .metric,
    values_from = .estimate
  ) %>%
  kable(
    digits = 3,
    booktabs = TRUE,
    align = "c",
    table.attr = 'data-quarto-disable-processing="true"'
  ) %>%
  kable_classic(
    full_width = FALSE
  )
```



































































